---
title: "NYPD Shooting Data Analysis"
author: "RT"
date: "2023-04-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing Data

 I stated importing the NYPD Shooting data from <https://catalog.data.gov/dataset/nypd-shooting-incident-data-historic>. So I created a "url" variable to read the data the website has.

```{r get_data}
## Get current Data in the "download" button of the website
url_2 <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
```
Let's read in the data and see what we have.

```{r import_data}
data_shooting <- read.csv(url_2)
```
After looking at the data_shooting, I would like to tidy the data set and remove the columns that I am not interested and rename other.
Also, I would like to see the range of date in my data set to start my analysis.
```{r tidy_data_shooting}
```
data_shooting <- data_shooting %>% select(-c(INCIDENT_KEY,OCCUR_TIME,JURISDICTION_CODE,LOCATION_DESC,STATISTICAL_MURDER_FLAG,PERP_AGE_GROUP,PERP_SEX,PERP_RACE,X_COORD_CD,Y_COORD_CD,Latitude,Longitude,Lon_Lat)) %>% rename(CITY = BORO, CITY AREA = PRECINCT)

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
